---
title: "Getting started with mverse!"
output: learnr::tutorial
runtime: shiny_prerendered
description: "Welcome to mverse!"
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggplot2)
# install.packages("remotes")
remotes::install_github("rstudio/gradethis")
gradethis::gradethis_setup()
#install.packages("tidyverse")

```

## Installation 

Install the `mverse` package from CRAN. You may have to set a mirror using `repos = "http://cran.us.r-project.org"`. 

```{r cranInstall, include=FALSE}
install.packages("mverse", repos = "http://cran.us.r-project.org")
```

```{r cranInstallExc, exercise=TRUE}

```

```{r cranInstallExc-hint}
install.packages( # insert more code here 
  )
```

```{r cranInstallExc-solution}
install.packages("mverse", repos = "http://cran.us.r-project.org")
```


## Warm-Up 1: Transform and Summarise One Numeric Column

Suppose that we have a column `col1` that we wish to transform in three different ways and compute the five number summary of the column after the transformations.

```{r ex1Installation, message=FALSE, exercise=TRUE, exercise.setup="cranInstall"}
library(tidyverse)
library(mverse)
set.seed(6) # For reproducibility
df <- tibble(col1 = rnorm(5, 0, 1), col2 = col1 + runif(5))
df
```

### Step 1: `create_multiverse` of the data frame

Create a multiverse object (`mv`) from `df` using `create_multiverse()`. 

```{r ex1step1, exercise.setup = "ex1Installation", include=FALSE}
mv <- create_multiverse(df)
```

```{r ex1step1Exc, exercise=TRUE, exercise.setup = "ex1Installation"}

```

```{r ex1step1Exc-solution}
mv <- create_multiverse(df)
```


### Step 2: `mutate_branch` to transform `col1`

Next, we will apply three different transformations to `col1`. Each transformation will create a different _branch_, and each branch corresponds to a universe. A branch in `mverse` refers to different modelling or data wrangling decisions. For example, a mutate branch - analogous to `mutate` method in `tidyverse`'s data manipulation grammar - lets you define a set of options for defining a new column in your dataset. 

You can create a mutate branch with `mutate_branch()`. The syntax for defining the options inside `mutate_branch()` follows the `tidyverse`'s grammar as well. 

In this case, Branch `col1` will have no transformation on `col1`, Branch `col1_t1` will take the log of absolute value of `col1 + 1`, and Branch `col1_t2` will take the absolute value of `col1`. 

Modify the code below to reflect the described transformations. 

```{r ex1step2, exercise.setup = "ex1step1", include=FALSE}
transformation <- mutate_branch(col1 = col1,
                                       col1_t1 = log(abs(col1 + 1)),
                                       col1_t2 = abs(col1))

```

```{r ex1step2Exc,  exercise=TRUE, exercise.setup = "ex1step1", exercise.blanks = TRUE}
transformation <- mutate_branch(col1 = ____,
                                       col1_t1 = ____,
                                       col1_t2 = ____)
```

```{r ex1step2Exc-hint-1, exercise.blanks = TRUE}
transformation <- mutate_branch(col1 = col1,
                                       col1_t1 = ____,
                                       col1_t2 = ____)
```

```{r ex1step2Exc-hint-2, exercise.blanks = TRUE}
transformation <- mutate_branch(col1 = col1,
                                       col1_t1 = log(abs(col1 + 1)),
                                       col1_t2 = ____)
```

```{r ex1step2Exc-hint-3}
transformation <- mutate_branch(col1 = col1,
                                       col1_t1 = log(abs(col1 + 1)),
                                       col1_t2 = abs(col1))
```


### Step 3: `add_mutate_branch` to `mv`

In this step, we will add the mutated branches (`transformation`) to our multiverse object, `mv`. Add the mutated branches to `mv` using `add_mutate_branch()`.

```{r ex1step3, exercise.setup = "ex1step2", include=FALSE}
mv <- mv %>% add_mutate_branch(transformation)
```

```{r ex1step3Exc, exercise=TRUE, exercise.setup = "ex1step2", excercise.blanks = TRUE}
mv <- mv %>% ____
```

```{r ex1step3Exc-solution}
mv <- mv %>% add_mutate_branch(transformation)
```

Adding a branch to an `mverse` object multiplies the number of environments defined inside the object so that the environments capture all unique analysis paths. Without any branches, a `mverse` object has a single environment. We call these environments _universes_. For example, adding the `transformation` mutated branch to `mv` results in $1 \times 3 = 3$ universes inside `mv`. In each universe, the analysis dataset now has a new column named `transformation` -- the name of the mutate branch object. 

You can verify that the mutated branch was added to `mv` by calling `summary` on the object. Run the code below to verify that there are three universes in this multiverse, and that each universe corresponds to one of the transformations on `col1`. 

```{r ex1step3B, exercise=TRUE, exercise.setup = "ex1step3"}
summary(mv)
```


### Step 4: `execute_multiverse` to execute the transformations

Now we will execute the analysis across the entire multiverse. Apply `execute_multiverse()` to `mv`, and save it to a multiverse object also called `mv`. 


```{r ex1step4, exercise.setup = "ex1step3", include=FALSE }
mv <- execute_multiverse(mv)
```

```{r ex1step4Exc, exercise=TRUE, exercise.setup = "ex1step3" , excercise.blanks = TRUE}
mv <- ____
```

```{r ex1step4Exc-solution}
mv <- execute_multiverse(mv)
```

### Step 5: Extract Transformed Values from `mv`

To extract the transformed values from `mv`, we can use the `extract` function. This function returns the transformed values along with their respective branch. We can then add this information to `df_transformed`. 

```{r ex1step5,exercise=TRUE, exercise.setup = "ex1step4" }
df_transformed <- extract(mv)
```

Use the `head()` function to explore `df_transformed`. 

```{r ex1step5B,exercise=TRUE, exercise.setup="ex1step5"}
```

```{r ex1step5B-solution}
df_transformed %>% head()
```

### Step 6: use `tidyverse` to compute the summary and plot the distribution of each transformation (universe)

We are now ready to compare the results from the different universes. Notice how different the mean, standard deviation, median, and IQR are between the universes.

We can also explore the universes visually using histograms. Feel free to change the number of bins to further investigate the shape of the distributions. 

```{r ex1step6A,exercise=TRUE, exercise.setup = "ex1step5"}
df_transformed %>% 
  group_by(transformation_branch_branch) %>% 
  summarise(n = n(), 
            mean = mean(transformation_branch), 
            sd = sd(transformation_branch),
            median = median(transformation_branch),
            IQR = IQR(transformation_branch))
 df_transformed %>% 
          ggplot(aes(x = transformation_branch)) + geom_histogram(bins = 3) +
          facet_wrap(vars(transformation_branch_branch))
```

## Warm-Up 2: Using `mverse` to Fit Three Simple Linear Regression of a Transformed Column

In this example, we again compare the three different transformations of `col1`. However, we will now use the three transformations to conduct simple linear regressions to predict `col2`. As such, this multiverse will have three universes -- one for each transformation of `col1`. 

### Step 1: `create_multiverse` of the data frame

Create a multiverse object named `mv1` from `df`. 

```{r ex2step1, exercise.setup = "ex1Installation", include=FALSE}
mv1 <- create_multiverse(df)
```

```{r ex2step1Exc, exercise=TRUE, exercise.setup = "ex1Installation"}

```

<div id="ex2step1Exc-hint">
**Hint:** You may want to use the `create_mutliverse()` function.
</div>


```{r ex2step1Exc-solution}
mv1 <- create_multiverse(df)
```

### Step 2: Create `formula_branch` of the linear regression models

In this scenario, each branch/universe corresponds to one of the three linear regressions using some transformation of `col1` to predict `col2`. In this step, we will create the three branches. 

`mverse` has a function specifically for adding formula branches called `formula_branch()`, which works similarly to the `mutate_branch()` function we used in Exercise 1. Run the code chunk below to create the `formulas` object. 

```{r ex2step2, exercise.setup = "ex2step1", exercise=TRUE}
formulas <- formula_branch(col2 ~ col1, 
                           col2 ~ log(abs(col1 + 1)), 
                           col2 ~ abs(col1))
```

### Step 3: `add_formula_branch` to multiverse of data frame

We will now add the branch `formulas` to our multiverse `mv1`. Using the function `add_formula_branch()`, add `formulas` to `mv1` and save the new multiverse object under the same name. 


```{r ex2step3, exercise.setup = "ex2step2", include=FALSE}
mv1 <- mv1 %>% add_formula_branch(formulas)
```

```{r ex2step3Exc, exercise=TRUE, exercise.setup = "ex2step2", excercise.blanks = TRUE}
mv1 <- mv1 %>% ____
```

<div id="ex2step3Exc-hint">
**Hint:** You may want to use the `add_formula_branch()` function.
</div>

```{r ex2step3Exc-solution}
mv1 <- mv1 %>% add_formula_branch(formulas)
```

### Step 4: `lm_mverse` to compute linear regression models across the multiverse

The `mverse` function `lm_mverse()` computes simple linear regression models across the entire multiverse object. Run the code chunk below to conduct this process. 

```{r ex2step4Exc, exercise=TRUE, exercise.setup = "ex2step3"}
lm_mverse(mv1)
```

### Step 5: `lm_mverse` to compute linear regression models across the multiverse

Write a line of code to view a summary of `mv1`, which now contains information about the linear regressions preformed.  

```{r ex2step5, exercise.setup = "ex2step4Exc", include=FALSE}
summary(mv1)
```

```{r ex2step5Exc, exercise=TRUE, exercise.setup = "ex2step4Exc"}

```

```{r ex2step5Exc-hint-1}
summary( #insert multiverse object here
  )
```

```{r ex2step5Exc-solution}
summary(mv1)
```

What do you notice about the predicted intercepts and coefficients for each of the branches? Are they similar to each other, or are they different? How does standard error vary across the models? Are the predictions significant? 

### Step 6: Comparing `mverse` to `tidyverse`

Given that we are only considering 3 branches, we can easily compare the results of the `mverse` analysis to what would be produced using `tidyverse` and base R. 

Using `tidyverse`, we will create a list of model formulas and map the list to the linear regression function `lm()`. Run the code below to conduct this process.

```{r ex2step6A, exercise=TRUE, exercise.setup = "ex2step5" }
mod1 <- formula(col2 ~ col1)
mod2 <- formula(col2 ~ log(abs(col1 + 1)))
mod3 <- formula(col2 ~ abs(col1))
models <- list(mod1, mod2, mod3)
models %>% 
  map(lm, data = df) %>% 
  map(broom::tidy) %>% 
  bind_rows()
```

Using base R, we will use `lapply` instead of mapping. Run the code below to conduct this process.

```{r ex2step6B, exercise=TRUE, exercise.setup = "ex2step5" }
modfit <- lapply(models, function(x) lm(x, data = df))
lapply(modfit, function(x) summary(x)[4])
```

Compare the results using `tidyverse` and base R to the multiverse summary from Step 5. The results should be nearly identical. 

You might be wondering why you need the `mverse` package at all if the analysis can be conducted using tools you already know. TODO: COME BACK TO THIS

## Exercise: Are Soccer Referees Biased?

```{r soccerLoad, include=FALSE}
library(mverse)
soccer_bias <- mverse::soccer[!is.na(mverse::soccer$rater1) & !is.na(mverse::soccer$rater2), 
                              c('playerShort', 'rater1', 'rater2')] # include only those with both ratings
soccer_bias <- unique(soccer_bias) # include only unique records for players
```


In this example, we use a real dataset to demonstrate how `mverse` allows to easily compare multiple definitions for a column, as well as the results from these different definitions. This dataset is included in the `mverse` package under the name `soccer`. 

The data comes from @datasrc and contains `r format(nrow(mverse::soccer), big.mark = ",")` rows of player-referee pairs. For each player, two independent raters coded their skin tones on a 5-point scale ranging from _very light skin_ (`0.0`) to _very dark skin_ (`1.0`). For the purpose of demonstration, we only use a unique record per player and consider only those with both ratings. 

We aim to combine the two ratings into one measure, but there are many ways to do so. In this example, we will explore how different methods of combining the ratings impact the distributions of the results. 

We would like to study the distribution of the player skin tones but the two independent rating do not always match. To combine the two ratings, we may choose to consider the following options:

1.  the mean of the two ratings
2.  the darker rating (closer to 1.0) of the two ratings
3.  the lighter rating  (closer to 0.0) of the two
4.  the first rating only
5.  the second rating only

First, we will load in the data. Run the chunk below to do so. 

```{r soccerLoadExc, exercise=TRUE,  exercise.setup = "cranInstall", message=FALSE}
library(mverse)
soccer_bias <- mverse::soccer[!is.na(mverse::soccer$rater1) & !is.na(mverse::soccer$rater2), 
                              c('playerShort', 'rater1', 'rater2')] # include only those with both ratings
soccer_bias <- unique(soccer_bias) # include only unique records for players
```


Now explore the dataset. Try to view the first six rows. 

```{r head, exercise=TRUE,  exercise.setup = "soccerLoadExc", message=FALSE}

```

<div id="head-hint">
  **Hint:** You may want to use the `head()` function.
</div>
  
```{r head-solution}
head(soccer_bias)
```

## Analysis using Base R and `Tidyverse`

Let's first consider how you might study the five options using R using `tidyverse` and base R, without `mverse`. First, we define the five rating combinations as separate variables in R.

The ratings for each player can be accessed with `soccer_bias$rater1` and `soccer_bias$rater2`. The first combination, the mean of the two ratings, has been written for you. Calculate the remaining four options according to the list above. You may want to make use of the `ifelse()` function. 

```{r base_r, exercise.setup = "soccerLoadExc", include=FALSE}
skin_option_1 <- (soccer_bias$rater1 + soccer_bias$rater2) / 2 
skin_option_2 <- ifelse(soccer_bias$rater1 > soccer_bias$rater2, soccer_bias$rater1, soccer_bias$rater2)
skin_option_3 <- ifelse(soccer_bias$rater1 < soccer_bias$rater2, soccer_bias$rater1, soccer_bias$rater2)
skin_option_4 <- soccer_bias$rater1    
skin_option_5 <- soccer_bias$rater2    
```

```{r base_rExc, exercise = TRUE, exercise.setup = "soccerLoadExc", excercise.blanks = TRUE}
skin_option_1 <- (soccer_bias$rater1 + soccer_bias$rater2) / 2 
skin_option_2 <- ____
skin_option_3 <- ____
skin_option_4 <- ____
skin_option_5 <- ____   
```

```{r base_rExc-solution}
skin_option_1 <- (soccer_bias$rater1 + soccer_bias$rater2) / 2 
skin_option_2 <- ifelse(soccer_bias$rater1 > soccer_bias$rater2, soccer_bias$rater1, soccer_bias$rater2)
skin_option_3 <- ifelse(soccer_bias$rater1 < soccer_bias$rater2, soccer_bias$rater1, soccer_bias$rater2)
skin_option_4 <- soccer_bias$rater1    
skin_option_5 <- soccer_bias$rater2    
```

We can plot a histogram to study the distribution of the resulting skin tone value for each option. Run the chunk below to view the histogram for the first option (`skin_option_1`).

```{r hist_base, exercise.setup = "base_r", exercise=TRUE}
library(ggplot2)
ggplot(mapping=aes(x = skin_option_1)) +
  geom_histogram(breaks = seq(0,1,0.2),
                 colour = 'white') +
  labs(title = 'Histogram of player skin tones (Option 1: Mean).',
       x = 'Skin Tone', y = 'Count')
```

For the remaining four options, we can repeat the step above to examine the distributions, or create a new data frame combining all five options to use in a ggplot as shown below. In both cases, users need to take care of plotting all five manually. Run the code below to view all five histograms. 

```{r hist_base_overlaid, exercise.setup = "base_r", exercise = TRUE}
skin_option_all <- data.frame(
  x = c(skin_option_1, skin_option_2, skin_option_3, skin_option_4, skin_option_5),
  Option = rep(c(
    'Option 1: Mean', 'Option 2: Max', 'Option 3: Min', 
    'Option 4: Rater 1', 'Option 5: Rater 2'), each = nrow(df)
    )
  )
ggplot(data = skin_option_all) +
  geom_histogram(aes(x = x), binwidth = 0.1) +
  labs(title = 'Histogram of player skin tones for each option.',
       x = 'Skin Tone', y = 'Count') +
  facet_wrap(. ~ Option)
```

## Analysis Using `mverse`

Perhaps using `mverse` requires less manual coding. We will now use this package to create the five options above. 

### Branching Using `mverse`

First, define an `mverse` object with the dataset.

```{r mvCreate, exercise.setup = "soccerLoad", include=FALSE, exercise = TRUE}
soccer_bias_mv <- create_multiverse(soccer_bias)
```

```{r mvCreateExc, exercise.setup = "soccerLoad", exercise = TRUE, exercise.blanks = TRUE}
soccer_bias_mv <- ____
```

<div id="mvCreateExc-hint">
**Hint:** You may want to use the `create_multiverse()` function.
</div>

```{r mvCreateExc-solution}
soccer_bias_mv <- create_multiverse(soccer_bias)
```

We can then use the `mutate_branch()` function to create the five different branches. In the chunk below, use this function and the the list of combinations to define the five branches.

```{r mutate_branch, exercise=TRUE, exercise.setup="mvCreate", include=FALSE}
skin_tone <- mutate_branch(
  (rater1 + rater2)/2,
  ifelse(rater1 > rater2, rater1, rater2),
  ifelse(rater1 < rater2, rater1, rater2),
  rater1,
  rater2
)
```

```{r mutate_branchExc, exercise=TRUE, exercise.setup="mvCreate", exercise.blanks=TRUE}
skin_tone <- ____
```

```{r mutate_branchExc-hint}
skin_tone <- mutate_branch(
  # calculation for average
  ,
  # calculation for max
  , 
  # calculation for min
  , 
  # rater1
  , 
  # rater 2
)
```

```{r mutate_branchExc-solution}
skin_tone <- mutate_branch(
  (rater1 + rater2)/2,
  ifelse(rater1 > rater2, rater1, rater2),
  ifelse(rater1 < rater2, rater1, rater2),
  rater1,
  rater2
)
```

Now, add the newly defined mutate branch to the `mv` object using `add_mutate_branch()`. 

```{r add_vb, exercise=TRUE, exercise.setup="mutate_branch", include=FALSE}
soccer_bias_mv <- soccer_bias_mv %>% add_mutate_branch(skin_tone)
```

```{r add_vbExc, exercise=TRUE, exercise.setup = "mutate_branch", excercise.blanks = TRUE}
soccer_bias_mv <- soccer_bias_mv %>% ____
```

<div id="add_vbExc-hint">
**Hint:** You may want to use the `add_mutate_branch()` function.
</div>

```{r add_vbExc-solution}
soccer_bias_mv <- soccer_bias_mv %>% add_mutate_branch(skin_tone)
```

Recall that adding a branch to an `mverse` object multiplies the existing environments within the object by the number of analysis paths within the new branch. Without any branches, the `soccer_bias_mv` object has a single environment -- a single universe. Because there are 5 analysis paths defined in `skin_tone`, we now have $1 \times 5 = 5$ environments. Our multiverse has expanded to encompass these 5 different universes. 

To verify that the mutated branch was added, call `summary()` on the `soccer_bias_mv` object. You should see a multiverse table that lists all universes, with the branches as columns and corresponding options as values.

```{r summaryExc, exercise=TRUE, exercise.setup = "add_vb"}
summary(soccer_bias_mv)
```

At this point, the values of the new column `skin_tone` are only populated in the first universe. To populate the values for all universes, we call `execute_multiverse()` on our multiverse object. Write the code below to conduct that process. 

```{r exec, exercise.setup="add_vb", exercise=TRUE, include=FALSE}
execute_multiverse(soccer_bias_mv)
```

```{r execExc, exercise.setup="add_vb", exercise=TRUE}

```

```{r execExc-solution}
execute_multiverse(soccer_bias_mv)
```



rmarkdown::run("gettingStartedHard.Rmd")
