---
title: "GLM modelling with mverse"
output: learnr::tutorial
runtime: shiny_prerendered
description: "Welcome to mverse!"
references:
  - id: original
    title: "Female hurricanes are deadlier than male hurricanes"
    type: article-journal
    issued:
      year: 2014
      month: 6
      day: 17
    URL: https://doi.org/10.1073/pnas.1402786111
    DOI: 10.1073/pnas.1402786111
    journal: Proceeedings fo the National Academiy of Sciences of the United States of America
    volume: 111
    issue: 24
    page: 8782-8787
    author:
    - given: Kiju
      family: Jung
    - given: Sharon
      family: Shavitt
    - given: Madhu
      family: Viswanathan
    - given: "Joseph M."
      family: Hilbe
  - id: multiverse
    title: "multiverse: R package for creating explorable multiverse analysis"
    type: entry
    URL: https://mucollective.github.io/multiverse/
    accessed:
      year: 2020
    author:
    - given: Abhraneel
      family: Sarma
    - given: Matthew
      family: Kay
  - id: boba
    title: "Boba: Authoring and Visualizing Multiverse Analyses"
    type: entry
    URL: https://arxiv.org/abs/2007.05551
    issued:
      year: 2020
      month: 7
      day: 30
    accessed:
      year: 2020
    author:
    - given: Yang
      family: Liu
    - given: Alex
      family: Kale
    - given: Tim
      family: Althoff
    - given: Jeffrey
      family: Heer
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(learnr)
library(tidyverse)
library(ggplot2)
library(mverse)

# install.packages("remotes")
remotes::install_github("rstudio/gradethis")
gradethis::gradethis_setup()
#install.packages("tidyverse")

```

## Introduction to GLMs in `mverse`


This tutorial aims to introduce the workflow of a multiverse analysis with generalized linear modelling using `mverse`.

The typical workflow of a multiverse analysis with `mverse` is

1.  Initialize a `multiverse` object with the dataset.
2.  Define all the different data analyses (i.e., analytical decisions) as *branches*.
3.  Add defined *branches* into the `multiverse` object.
4.  Run models, hypothesis tests, and plots.

## Exploring The Severity Of Feminine-named Versus Masculine-named Hurricanes

In this example, we use a real dataset to demonstrate how `mverse` allows us to easily compare variations of generalized linear models (GLMs). Specifically, we will explore the severity of feminine- versus masculine-named hurricanes. This dataset is included in the `mverse` package under the name `hurricane`. The data comes from @original. `mverse` has already been loaded. 

Below is a glimpse of the `hurricane` data.

```{r viewData, exercise=TRUE, exercise.setup="setup", exercise.eval=TRUE}
glimpse(hurricane)
```

To start a multiverse analysis, first use `create_multiverse` to create an `mverse` object with `hurricane`. Name this multiverse object `hurricane_mv`. 

```{r createMV, exercise.setup = "setup", include=FALSE}
hurricane_mv <- create_multiverse(hurricane)
```

```{r createMVExc, exercise = TRUE, exercise.setup = "setup"}

```

```{r createMVExc-hint}
hurricane_mv <- create_multiverse( #data
  )
```

```{r createMVExc-solution}
hurricane_mv <- create_multiverse(hurricane)
```


Remember that at this point, the multiverse contains only _one_ environment (the default). 

### Define Branches in the Hurricane Multiverse

Each branch defines a different statistical analysis by using a subset of data, transforming columns, or a statistical model. Each combination of these branches defines a "universe", or analysis path. Branches are defined using `X_branch(...)`, where `...` are expressions that define a data wrangling or modelling option/analytic decision that you wish to explore -- such as excluding certain hurricane names from an analysis, deriving new variables for analysis (mutating), or using different models. Once branches are defined, we can look at the impact of the combination of these decisions.

First we will explore branches that manipulate the data itself, and then we will add branches that apply different statistical models. 


#### Branches for Data Manipulation

The `mverse` function `filter_branch` takes logical predicates, and finds the observations where the condition is `TRUE`. It can also be used to exclude outliers. 

An important variable in the `hurricane` dataset is `alldeaths`, which indicates the number of deaths caused by a particular hurricane. Below is the distribution of `alldeaths`. 

```{r alldeathsHist, exercise.eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(mverse)
library(tidyverse)
library(dplyr)
hurricane %>% 
  ggplot(aes(alldeaths)) + 
  geom_histogram(bins = 25) + 
  stat_bin(aes(label = ..count..), bins = 25, geom = "text", vjust = -.7, size = 2)
```

It seems like there are two extreme outliers -- let's find out which hurricanes those are. 

```{r outliers, exercise=TRUE, exercise.setup="createMV", exercise.eval=TRUE}
hurricane %>% 
  filter(alldeaths > median(alldeaths)) %>% 
  arrange(desc(alldeaths)) %>% 
  select(Name, alldeaths) %>% 
  head()

```

It seems like hurricanes Katrina and Audrey have the most extreme death counts. 

There are a few different ways to address outliers in data wrangling. Firstly, we can simply include them in our analysis. We can also just remove the most extreme observation -- in this case, Hurricane Katrina. Finally, we can exclude the two most extreme observations -- both Hurricane Katrina and Hurricane Audrey. We can use `filter_branch` to consider each of these options in our multiverse.

The code below creates a branch that considers all three outlier-handling options. 

```{r filter_branch, exercise=TRUE, exercise.eval=TRUE, exercise.setup="createMV"}
death_outliers <- filter_branch(obs_all = TRUE, 
                                obs_noKat = Name != 'Katrina',
                                obs_noKatAud = !(Name %in% c('Katrina', 'Audrey')))
```

Now, add this branch to `hurricane_mv` using the `add_filter_branch()` function. 

```{r addBranch, exercise.setup="filter_branch", include=FALSE}
hurricane_mv <- hurricane_mv %>% add_filter_branch(death_outliers)
```

```{r addBranchExc,  exercise=TRUE, exercise.setup = "filter_branch", exercise.blanks = TRUE}
hurricane_mv <- hurricane_mv %>% ____
```


```{r addBranchExc-solution}
hurricane_mv <- hurricane_mv %>% add_filter_branch(death_outliers)
```

Now, call the `summary()` function on the mutliverse object `hurricane_mv` to ensure that there are three universes, and that each universe corresponds to an outlier-handling option. 

```{r summary, exercise=TRUE, exercise.setup="addBranch"}

```

```{r summary-solution}
summary(hurricane_mv)
```

Now suppose we want to consider two different definitions of femininity as provided in the dataset. `Gender_MF` is a binary classification the gender of the name of a hurricane, and `MasFem` is a continuous rating of femininity for hurricane names. We can use `mutate_branch()` to create a branch that details different transformations or definitions of the same variable. 

The code below creates the `femininity` branch. 

```{r mutate_branch, exercise=TRUE, exercise.eval=TRUE, exercise.setup="addBranch"}
femininity <- mutate_branch(binary_gender = Gender_MF, 
                            cts_gender = MasFem)
```

Suppose we also want to include normalized damage (NDAM) in our analysis. Below is a histogram of the `NDAM` variable.

```{r NDAMhist, exercise=TRUE, exercise.setup="mutate_branch", exercise.eval=TRUE}
hurricane %>% 
  ggplot(aes(NDAM)) + 
  geom_histogram(bins = 25) + 
  stat_bin(aes(label = ..count..), bins = 25, geom = "text", vjust = -.7, size = 2)
```

Evidently, the NDAM variable is fairly right-skewed. Perhaps we want to consider log-transforming this variable to reduce the skewness. `mutate_branch()` can help us do so. 

Following the creation of the `femininity` branch above, create a branch named `damage` with two options: 

* `damage_orig`, which the same as the raw `NDAM` value in the dataset
* `damage_log`, which is the log of the `NDAM` value in the dataset

```{r mutate_branchDam,  exercise.setup="addBranch", include=FALSE}
damage <- mutate_branch(damage_orig = NDAM, 
                        damage_log = log(NDAM))
```

```{r mutate_branchDamExc, exercise=TRUE, exercise.setup="addBranchExc"}

```

```{r mutate_branchDamExc-hint-1}
damage <- 
```

```{r mutate_branchDamExc-hint-2}
damage <- mutate_branch( # create damage_orig,
                         # create damage_log
  )
```

```{r mutate_branchDamExc-hint-3}
damage <- mutate_branch( damage_orig = NDAM,
                         # create damage_log
  )
```

```{r mutate_branchDamExc-solution}
damage <- mutate_branch( damage_orig = NDAM, 
                        damage_log = log(NDAM))
```

Now, add both the `femininity` and `damage` branches to our multiverse object. 

```{r addBranches, exercise.setup="mutate_branchDam", include=FALSE}
hurricane_mv <- hurricane_mv %>% add_mutate_branch(femininity, damage)
```

```{r addBranchesExc, exercise=TRUE, exercise.setup="mutate_branchDam", exercise.blanks=TRUE }
hurricane_mv <- hurricane_mv %>% ____
```

```{r addBranchesExc-hint-1}
hurricane_mv <- hurricane_mv %>% add_mutate_branch( # branches
  )
```

```{r addBranchesExc-solution}
hurricane_mv <- hurricane_mv %>% add_mutate_branch(femininity, damage)
```

Now we have added all three branches to our universe: the outlier branch, the femininity branch, and the damage branch. Remember that our multiverse began with just one universe -- how many should it have now? 

```{r universes, echo=FALSE}
library(learnr)
question("How many universes should be in our multiverse?",
  answer("1"),
  answer("7"),
  answer("12", correct = TRUE),
  answer("20"),
  incorrect = "Remember that the outlier branch has 3 options, the femininity branch has 2 options, and the damage branch has 2 options. Each combination of these options corresponds to a universe. How many different combinations are there? ",
      allow_retry = TRUE,
  post_message = "You got it! Since the outlier branch has 3 options, the femininity branch has 2 options, and the damage branch has 2 options, we get 3 x 2 x 2 = 12 universes in total."

)
```

To check your understanding, call `summary()` on our multiverse to see how many universes it contains. 

```{r summaryExc, exercise.setup = 'addBranches', exercise=TRUE}

```

```{r summaryExc-solution}
summary(hurricane_mv)
```






rmarkdown::run("GLMmodelling.Rmd")
