---
title: "GLM modelling with mverse"
output: learnr::tutorial
runtime: shiny_prerendered
description: "Welcome to mverse!"
references:
  - id: original
    title: "Female hurricanes are deadlier than male hurricanes"
    type: article-journal
    issued:
      year: 2014
      month: 6
      day: 17
    URL: https://doi.org/10.1073/pnas.1402786111
    DOI: 10.1073/pnas.1402786111
    journal: Proceeedings fo the National Academiy of Sciences of the United States of America
    volume: 111
    issue: 24
    page: 8782-8787
    author:
    - given: Kiju
      family: Jung
    - given: Sharon
      family: Shavitt
    - given: Madhu
      family: Viswanathan
    - given: "Joseph M."
      family: Hilbe
  - id: multiverse
    title: "multiverse: R package for creating explorable multiverse analysis"
    type: entry
    URL: https://mucollective.github.io/multiverse/
    accessed:
      year: 2020
    author:
    - given: Abhraneel
      family: Sarma
    - given: Matthew
      family: Kay
  - id: boba
    title: "Boba: Authoring and Visualizing Multiverse Analyses"
    type: entry
    URL: https://arxiv.org/abs/2007.05551
    issued:
      year: 2020
      month: 7
      day: 30
    accessed:
      year: 2020
    author:
    - given: Yang
      family: Liu
    - given: Alex
      family: Kale
    - given: Tim
      family: Althoff
    - given: Jeffrey
      family: Heer
  - id: spec
    title: "Specification curve analysis"
    type: article-journal
    URL: https://www.nature.com/articles/s41562-020-0912-z
    issued:
      year: 2020
      month: 7
      day: 27
    accessed:
      year: 2023
    author:
    - given: Uri
      family: Simonsohn
    - given: Joseph
      family: Simmons
    - given: Leif D. 
      family: Nelson
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(learnr)
library(tidyverse)
library(ggplot2)
library(mverse)
#install.packages("tidyverse")
```

## Introduction to GLMs in `mverse`


This tutorial aims to introduce the workflow of a multiverse analysis with generalized linear modelling using `mverse`.

The typical workflow of a multiverse analysis with `mverse` is

1.  Initialize a `multiverse` object with the dataset.
2.  Define all the different data analyses (i.e., analytical decisions) as *branches*.
3.  Add defined *branches* into the `multiverse` object.
4.  Run models, hypothesis tests, and plots.

## Exploring The Severity Of Feminine-named Versus Masculine-named Hurricanes

In this example, we use a real dataset to demonstrate how `mverse` allows us to easily compare variations of generalized linear models (GLMs). Specifically, we will explore the severity of feminine- versus masculine-named hurricanes. This dataset is included in the `mverse` package under the name `hurricane`. The data comes from @original. `mverse` has already been loaded. 

Below is a glimpse of the `hurricane` data.

```{r viewData, exercise=TRUE, exercise.setup="setup", exercise.eval=TRUE}
glimpse(hurricane)
```

To start a multiverse analysis, first use `create_multiverse` to create an `mverse` object with `hurricane`. Name this multiverse object `hurricane_mv`. 

```{r createMV, exercise.setup = "setup", include=FALSE}
hurricane_mv <- create_multiverse(hurricane)
```

```{r createMVExc, exercise = TRUE, exercise.setup = "setup"}
```

```{r createMVExc-hint}
hurricane_mv <- create_multiverse( #data
  )
```

```{r createMVExc-solution}
hurricane_mv <- create_multiverse(hurricane)
```


Remember that at this point, the multiverse contains only _one_ environment (the default). 

### Define Branches in the Hurricane Multiverse

Each branch defines a different statistical analysis by using a subset of data, transforming columns, or a statistical model. Each combination of these branches defines a "universe", or analysis path. Branches are defined using `X_branch(...)`, where `...` are expressions that define a data wrangling or modelling option/analytic decision that you wish to explore -- such as excluding certain hurricane names from an analysis, deriving new variables for analysis (mutating), or using different models. Once branches are defined, we can look at the impact of the combination of these decisions.

First we will explore branches that manipulate the data itself, and then we will add branches that apply different statistical models. 


#### Branches for Data Manipulation

The `mverse` function `filter_branch` takes logical predicates, and finds the observations where the condition is `TRUE`. It can also be used to exclude outliers. 

An important variable in the `hurricane` dataset is `alldeaths`, which indicates the number of deaths caused by a particular hurricane. Below is the distribution of `alldeaths`. 

```{r alldeathsHist, exercise.eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(mverse)
library(tidyverse)
library(dplyr)
hurricane %>% 
  ggplot(aes(alldeaths)) + 
  geom_histogram(bins = 25) + 
  stat_bin(aes(label = ..count..), bins = 25, geom = "text", vjust = -.7, size = 2)
```

It seems like there are two extreme outliers -- let's find out which hurricanes those are. 

```{r outliers, exercise=TRUE, exercise.setup="createMV", exercise.eval=TRUE}
hurricane %>% 
  filter(alldeaths > median(alldeaths)) %>% 
  arrange(desc(alldeaths)) %>% 
  select(Name, alldeaths) %>% 
  head()
```

It seems like hurricanes Katrina and Audrey have the most extreme death counts. 

There are a few different ways to address outliers in data wrangling. Firstly, we can simply include them in our analysis. We can also just remove the most extreme observation -- in this case, Hurricane Katrina. Finally, we can exclude the two most extreme observations -- both Hurricane Katrina and Hurricane Audrey. We can use `filter_branch` to consider each of these options in our multiverse.

The code below creates a branch that considers all three outlier-handling options. 

```{r outlierBranch, exercise=TRUE, exercise.eval=TRUE, exercise.setup="createMV"}
death_outliers <- filter_branch(obs_all = TRUE, 
                                obs_noKat = Name != 'Katrina',
                                obs_noKatAud = !(Name %in% c('Katrina', 'Audrey')))
```

Now, add this branch to `hurricane_mv` using the `add_filter_branch()` function. 

```{r addOutlier, exercise.setup="outlierBranch", include=FALSE}
hurricane_mv <- hurricane_mv %>% add_filter_branch(death_outliers)
```

```{r addOutlierExc,  exercise=TRUE, exercise.setup = "outlierBranch", exercise.blanks = TRUE}
hurricane_mv <- hurricane_mv %>% ____
```


```{r addOutlierExc-solution}
hurricane_mv <- hurricane_mv %>% add_filter_branch(death_outliers)
```

Now, call the `summary()` function on the mutliverse object `hurricane_mv` to ensure that there are three universes, and that each universe corresponds to an outlier-handling option. 

```{r summary, exercise=TRUE, exercise.setup="addOutlier"}
```

```{r summary-solution}
summary(hurricane_mv)
```

Now suppose we want to consider two different definitions of femininity as provided in the dataset. `Gender_MF` is a binary classification the gender of the name of a hurricane, and `MasFem` is a continuous rating of femininity for hurricane names. We can use `mutate_branch()` to create a branch that details different transformations or definitions of the same variable. 

The code below creates the `femininity` branch. 

```{r femBranch, exercise=TRUE, exercise.eval=TRUE, exercise.setup="addOutlier"}
femininity <- mutate_branch(binary_gender = Gender_MF, 
                            cts_gender = MasFem)
```

Suppose we also want to include normalized damage (NDAM) in our analysis. Below is a histogram of the `NDAM` variable.

```{r NDAMhist, exercise=TRUE, exercise.setup="femBranch", exercise.eval=TRUE}
hurricane %>% 
  ggplot(aes(NDAM)) + 
  geom_histogram(bins = 25) + 
  stat_bin(aes(label = ..count..), bins = 25, geom = "text", vjust = -.7, size = 2)
```

Evidently, the NDAM variable is fairly right-skewed. Perhaps we want to consider log-transforming this variable to reduce the skewness. `mutate_branch()` can help us do so. 

Following the creation of the `femininity` branch above, create a branch named `damage` with two options: 

* `damage_orig`, which the same as the raw `NDAM` value in the dataset
* `damage_log`, which is the log of the `NDAM` value in the dataset

```{r damageBranch,  exercise.setup="femBranch", include=FALSE}
damage <- mutate_branch(damage_orig = NDAM, 
                        damage_log = log(NDAM))
```

```{r damageBranch-solution}
damage <- mutate_branch( damage_orig = NDAM, 
                        damage_log = log(NDAM))
```

Now, add both the `femininity` and `damage` branches to our multiverse object. 

```{r addBranches, exercise.setup="damageBranch", include=FALSE}
hurricane_mv <- hurricane_mv %>% add_mutate_branch(femininity, damage)
```

```{r addBranchesExc, exercise=TRUE, exercise.setup="damageBranch", exercise.blanks=TRUE }
hurricane_mv <- hurricane_mv %>% ____
```

```{r addBranchesExc-hint-1}
hurricane_mv <- hurricane_mv %>% add_mutate_branch( # branches
  )
```

```{r addBranchesExc-solution}
hurricane_mv <- hurricane_mv %>% add_mutate_branch(femininity, damage)
```

Now we have added all three branches to our universe: the outlier branch, the femininity branch, and the damage branch. Remember that our multiverse began with just one universe -- how many should it have now? 

```{r universes, echo=FALSE}
library(learnr)
question("How many universes should be in our multiverse?",
  answer("1"),
  answer("7"),
  answer("12", correct = TRUE),
  answer("20"),
  incorrect = "Remember that the outlier branch has 3 options, the femininity branch has 2 options, and the damage branch has 2 options. Each combination of these options corresponds to a universe. How many different combinations are there? ",
      allow_retry = TRUE,
  post_message = "You got it! Since the outlier branch has 3 options, the femininity branch has 2 options, and the damage branch has 2 options, we get 3 x 2 x 2 = 12 universes in total."
)
```

To check your understanding, call `summary()` on our multiverse to see how many universes it contains. 

```{r summaryExc, exercise.setup = 'addBranches', exercise=TRUE}
```

```{r summaryExc-solution}
summary(hurricane_mv)
```


#### Branches for GLM modelling

`mverse` can also define different `glm()` models as branches. The formula for a `glm` model (e.g., `y ~ x`) can be defined using `formula_branch()`, and `family_branch()` defines the member of the exponential family used via a `family` object.

We can create formulas using the branches above or simply use the columns in dataframe. Suppose we want to use `damage` and `femininity` to predict `alldeaths`. We can define a branch `models` using the `formula_branch()` function, and can add it to our multiverse using `add_formula_branch()`. 

```{r models, exercise=TRUE, exercise.setup="addBranches", exercise.eval=TRUE}
models <- formula_branch(alldeaths ~ damage + femininity)
hurricane_mv <- hurricane_mv %>% add_formula_branch(models)
```


Now we have added the `models` branch to our multiverse object. Notice that this branch only has one option -- in other words, all existing universes are prescribed the same formula. After adding the `models` branch, how many universes exist in our multiverse? 

```{r universesModels, echo=FALSE}
library(learnr)
question("How many universes should be in our multiverse?",
  answer("12", correct = TRUE),
  answer("24"),
  answer("144"),
  incorrect = "If there are 12 existing universes, and our new branch only has one option, how many combinations result? ",
      allow_retry = TRUE,
  post_message = "You got it! Since the multiverse had 12 universes, and the models branch has 1 option, we get 12 x 1 = 12 universes in total."
)
```

To check your understanding, call `summary()` on our multiverse to see how many universes it contains. 

```{r modelsSummary, exercise.setup = 'models', exercise=TRUE}
```

```{r modelsSummary-solution}
summary(hurricane_mv)
```

Finally, let's create a `distributions` branch that defines two different members of the Exponential family using `family_branch()`.

```{r familyBranch, exercise=TRUE, exercise.setup="models", exercise.eval=TRUE}
distributions <- family_branch(poisson, gaussian)
hurricane_mv <- hurricane_mv %>% add_family_branch(distributions)
```

After adding the `distributions` branch to our multiverse, how many universes should it contain? 

```{r universesDistributions, echo=FALSE}
library(learnr)
question("How many universes should be in our multiverse?",
  answer("12"),
  answer("24", correct = TRUE),
  answer("144"),
  incorrect = "If there are 12 existing universes, and our new branch has two options, how many combinations result? ",
      allow_retry = TRUE,
  post_message = "You got it! Since the multiverse had 12 universes, and the distribution branch has 2 options, we get 12 x 2 = 24 universes in total."
)
```

To check your understanding, call `summary()` on our multiverse to see how many universes it contains. 

```{r distributionsSummary, exercise.setup = 'familyBranch', exercise=TRUE}
```

```{r distributionsSummary-solution}
summary(hurricane_mv)
```

When dealing with multiverse problems, visualization can be helpful to understand the composition of each universe. The `multiverse_tree()` function can be used to visualize multiverses, whether you want to view the entire thing or just a subset of branches.

The tree below shows the entire `hurricane_mv` multiverse. Feel free to adjust `label_size` and `label_angle`. 

```{r treeAll, exercise=TRUE, exercise.eval=TRUE, exercise.setup="familyBranch", warning=FALSE}
multiverse_tree(hurricane_mv, label_size = 2, label_angle = 45, label=TRUE)
```

Suppose that you just want to view the `models` and `distribution` branches. Then, you can specify these branches as a vector argument to `branches`. 

```{r treeModDist, exercise=TRUE, exercise.eval=TRUE, exercise.setup="familyBranch", warning=FALSE}
multiverse_tree(hurricane_mv, label = TRUE, label_size = 4, label_angle = 45, branches = c("models", "distributions"))
```

Create a tree that just shows the `femininity` and `damage` branches. Use `label_size = 4` and `label_angle = 45`.

```{r treeFemDam, include=FALSE, exercise.setup="familyBranch", warning=FALSE}
multiverse_tree(hurricane_mv, label = TRUE, label_size = 4, label_angle = 45, branches = c("femininity", "damage"))
```

```{r treeFemDamExc, exercise=TRUE, exercise.setup="familyBranch"}
```

```{r treeFemDamExc-solution}
multiverse_tree(hurricane_mv, label = TRUE, label_size = 4, label_angle = 45, branches = c("femininity", "damage"))
```


Now we are ready to fit the GLM specified in the `models` to each universe. We can execute this process by calling the `glm_mverse()` function on our multiverse object. 

```{r glmExecute, exercise=TRUE, exercise.eval=TRUE, exercise.setup="treeFemDam"}
glm_mverse(hurricane_mv)
summary(hurricane_mv)
```

Be careful not to mistake the number of rows in the multiverse summary with the number of universes. We still only have 24 universes, but each universe now spreads across three rows -- each row corresponds to one of the model coefficients (intercept, coefficient for damage, coefficient for femininity).

Suppose you want to investigate how a particular model coefficient behaves across the different universe. The function `spec_curve()` can help you investigate these questions. 

Proposed by @spec, specification curves allow users to identify influential decisions, such as how a variable is defined, what kind of model is used, etc. By presenting coefficient estimates and standard errors in this format, users can identify possible sources of noise and bias in their analysis.

The specification curve for the coefficient of `femininity` is shown below.

```{r specFem, exercise=TRUE, exercise.eval=TRUE, exercise.setup="glmExecute"}
spec_curve(hurricane_mv, var = "femininity", color_order = TRUE)
```

What do you notice? 

<TODO: some kind of question here?>


#### Condition Branches 

Recall the distribution of `alldeaths`. 

```{r allDeathsHist2, exercise=TRUE, exercise.eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, exercise.setup="specFem"}
hurricane %>% 
  ggplot(aes(log(alldeaths+1))) + 
  geom_histogram(bins = 25) + 
  stat_bin(aes(label = ..count..), bins = 25, geom = "text", vjust = -.7, size = 2)
```

Even if we were to remove the two outliers (Hurricanes Katrina and Audrey), the assumption that `alldeaths` follows a Normal distribution is tenuous -- but this is an assumption for our GLM process to be valid. We can check this condition further by creating a quantile-quantile (QQ) plot of the variable.

```{r allDeathsQQ2, exercise=TRUE, exercise.eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, exercise.setup="specFem"}
hurricane %>% ggplot(aes(sample = alldeaths)) + stat_qq() + stat_qq_line()
```

Given the deviation of the dots (the data) from the solid line (a normal distribution), it is unlikely that `alldeaths` is Normally distributed.

However, transforming the `alldeaths` using $t(x)=\log(x+1)$ could result in a dependent variable that is closer to a Normal distribution. Note: we shift the variable over by one ($x+1$) because many hurricanes have 0 deaths, but $log(0)$ is undefined.  

Let's apply the transformation and see how the histogram and QQ plot changes. 

```{r allDeathsTransformedHist, exercise=TRUE, exercise.eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, exercise.setup="specFem"}
hurricane %>% 
  ggplot(aes(log(alldeaths+1))) + 
  geom_histogram(bins = 25) + 
  stat_bin(aes(label = ..count..), bins = 25, geom = "text", vjust = -.7, size = 2)
```

```{r allDeathsTransformedQQ, exercise=TRUE, exercise.eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, exercise.setup="specFem"}
hurricane %>% mutate(logd = log(alldeaths+1)) %>% ggplot(aes(sample = logd)) + stat_qq() + stat_qq_line()
```

While there is still some skewness/deviation, the transformation certainly helped bring the `alldeaths` distribution closer to Normal. 

Now, let's set up a similar multiverse analysis to the one above. 

Create the multiverse. 

```{r mv1, include=FALSE, exercise.setup="setup"}
hurricane_mv <- create_multiverse(hurricane)
```

```{r mv1Exc, exercise.setup="setup", exercise=TRUE}
```

```{r mv1Exc-hint}
hurricane_mv <- create_multiverse( #data
  )
```

```{r mv1Exc-solution}
hurricane_mv <- create_multiverse(hurricane)
```

Create a branch called `dep_var` that considers `alldeaths` as well as `log(alldeaths + 1)`. 

```{r mv2, include=FALSE, exercise.setup="mv1"}
dep_var <- mutate_branch(alldeaths, log(alldeaths + 1))
```

```{r mv2Exc, exercise.setup="mv1", exercise=TRUE}
```

```{r mv2Exc-hint}
dep_var <- mutate_branch( # insert options
  )
```

```{r mv2Exc-solution}
dep_var <- mutate_branch(alldeaths, log(alldeaths + 1))
```
Create a branch called `femininity` that considers two definitions of gender from the dataset: `Gender_MF` and `MasFem`. Hint: look at the previous example.

```{r mv3, include=FALSE, exercise.setup="mv2"}
femininity <- mutate_branch(binary_gender = Gender_MF, 
                            cts_gender = MasFem)
```

```{r mv3Exc, exercise.setup="mv2", exercise=TRUE}
```

```{r mv3Exc-hint}
femininity <- mutate_branch( # insert options
  )
```

```{r mv3Exc-solution}
femininity <- mutate_branch(binary_gender = Gender_MF, 
                            cts_gender = MasFem)
```

Create a branch called `damage` that considers `damage` as well as `log(damage)`.

```{r mv4, include=FALSE, exercise.setup="mv3"}
damage <- mutate_branch(damage_orig = NDAM, 
                        damage_log = log(NDAM))
```

```{r mv4Exc, exercise.setup="mv3", exercise=TRUE}
```

```{r mv4Exc-hint}
damage <- mutate_branch( # insert options
  )
```

```{r mv4Exc-solution}
damage <- mutate_branch(binary_gender = Gender_MF, 
                            cts_gender = MasFem)
```

Create a branch called `models` that defines the GLM for each universe: `dep_var ~ damage + femininity`. 


```{r mv5, include=FALSE, exercise.setup="mv4"}
models <- formula_branch(dep_var ~ damage + femininity)
```

```{r mv5Exc, exercise.setup="mv4", exercise=TRUE}
```

```{r mv5Exc-hint}
models <- formula_branch( # insert options
  )
```

```{r mv5Exc-solution}
models <- formula_branch(dep_var ~ damage + femininity)
```

Create a branch called distributions that considers two possible Exponential families, `poisson` and `gaussian`. 

```{r mv6, include=FALSE, exercise.setup="mv5"}
distributions <- family_branch(poisson, gaussian)
```

```{r mv6Exc, exercise.setup="mv5", exercise=TRUE}
```

```{r mv6Exc-hint}
distributions <- family_branch( # insert options
  )
```

```{r mv6Exc-solution}
distributions <- family_branch(poisson, gaussian)
```

Add these branches to the multiverse object. 

```{r mv7, include=FALSE, exercise.setup="mv6"}
hurricane_mv <- hurricane_mv %>% 
  add_mutate_branch(dep_var, femininity, damage) %>%
  add_formula_branch(models) %>% 
  add_family_branch(distributions)
```

```{r mv7Exc, exercise.setup="mv6", exercise=TRUE}
```

```{r mv7Exc-hint}
hurricane_mv <- hurricane_mv %>% 
 # add branches 
```

```{r mv7Exc-solution}
hurricane_mv <- hurricane_mv %>% 
  add_mutate_branch(dep_var, femininity, damage) %>%
  add_formula_branch(models) %>% 
  add_family_branch(distributions)
```

Use `multiverse_tree()` to display the multiverse tree of `dep_var` and `distributions`. Use `label_size=4` and `label_angle=45`. 

```{r mv8, include=FALSE, exercise.setup="mv7" }
multiverse_tree(hurricane_mv, label = TRUE, label_size = 4, label_angle = 45, branches = c("dep_var", "distributions"))
```

```{r mv8Exc, exercise.setup="mv7", exercise=TRUE }
```

```{r mv8Exc-hint}
multiverse_tree(hurricane_mv, label = TRUE, label_size = 4, label_angle = 45, branches = c("dep_var", "distributions"))
```

However, suppose we don't necessarily want to have all combinations of all the branches in our multiverse. Perhaps we only want to specify a Gaussain family to analyses using `log(alldeaths + 1)` and a Poisson family to analyses using `alldeaths`. In this case, we can make use of the `branch_condition()` method and use `add_branch_condition()` to add it to `hurricane_mv`.

```{r mv9, exercise=TRUE, exercise.setup="mv8", exercise.eval=TRUE}
match_poisson <- branch_condition(alldeaths, poisson)
```


Before adding the condition, we had a multiverse where there were 2 options for the dependent variabe, 2 options for femininity, 2 options for damage, 1 option for the model, and 2 options for the distribution family. After imposing this condition on our multiverse, how many universes should it contain? 

```{r universesGLM, echo=FALSE}
library(learnr)
question("How many universes should be in our multiverse?",
  answer("8", correct = TRUE),
  answer("12"),
  answer("16"),
  incorrect = "Before, each option for dependent variable was paired with each option distribution family. Now, each option for dependent variable can only be paired with one kind of distribution family. What effect does this have?",
      allow_retry = TRUE,
  post_message = "You got it! Before imposing the condition, we had 2 x 2 x 2 x 1 x 2 = 16 universes. However, when each option for the dependent variable can only be paired with one kind of distribution family, we are left with 2 x 2 x 2 x 1 x 1 = 8 universes."
)
```

The multiverse tree now shows that `log(alldeaths + 1)` will _only_ be modelled as Gaussian and `alldeaths` will _only_ be modelled as Poisson.

```{r mv10, excercise=TRUE, exercise.setup="mv9", exercise.eval=TRUE }
multiverse_tree(hurricane_mv, label_size=4, label_angle = 45, label = TRUE, c("dep_var", "distributions"))
```

Now, apply `glm_mverse()` to our multiverse object. 

```{r mv11, exercise=TRUE, exercise.setup="mv10", exercise.eval=TRUE}
glm_mverse(hurricane_mv)
summary(hurricane_mv)
```

A summary of the models on the `femininity` coefficients is shown in the specification curve.

```{r mv12, exercise=TRUE, exercise.setup="mv11", exercise.eval=TRUE}
spec_curve(hurricane_mv, var = "femininity", color_order = TRUE)
```
rmarkdown::run("GLMmodelling.Rmd")